Temperature Measurements in Optical Tweezer Experiments

### 2017-01-25 doc/draft2.tex
work work work

\newpage
\section{Motivation}
%introduction of the experiment, the problem and the idea
\section{The experiment}
The starting point of this thesis is an experiment conducted by Gieseler et al \cite{Gieseler2014}. It is an optical tweezer experiment, where the
motion of a glass nanoparticle in a laser trap was used to investigate the fluctuation theorem\cite{Crooks1999}.
\subsection{Experimental setup}
In the experiment, a silica nano particle with a radius of about 75 nm and mass of about $3 \times 10^{-18}$kg is trapped in a laser beam within a
vacuum chamber. The trapping of the silica nano particle (which will be referred to as \textit{glass particle}) is achieved by a gradient force of the
laser beam acting on the particle. The experimental setup is depicted in fig.~\ref{fig:setup}.\\
The particle fluctuates within the trap in all three spatial directions. These fluctuations can be approximated
such that they are decoupled, which means that they can be described by a 1-dimensional Langevin equation:
\begin{equation}
    \ddot{x} + \Gamma_0 \dot{x} + \Omega^2_0x = \frac 1 m \left(F_\text{fluct} + F_\text{ext}\right)
\end{equation}
%The experiment is set up as follows: a glass sphere with radius $r \approx 75$ nm and mass $m \approx 3 \times 10^{-18}$ kg is trapped in a laser beam
%in a vacuum chamber. The initial state of the glass particle is prepared by modulation of the laser, such that the oscillation of the glass particle
%is suppressed in every direction. This is done by measuring the oscillation of the particle in every direction (assuming that the oscillation in every
%direction is decoupled) and using this information as feedback for the laser. This experimental setup can be seen in figure \ref{fig:setup}.
\begin{figure}[H]
    \begin{center}
        \includegraphics[scale=0.3]{images/experimental_setup.jpg}
        \caption{Experimental setup of the optical tweezer experiment. A silica nano particle is trapped in a laser beam via gradient force in a
        vacuum. The feedback is used to cool down the particle and create a non-equilibrium steady state. In the first part of the experiment, the
    feedback is turned off and the motion of the particle towards an equilibrated state is observed. In the second part of the experiment, the steady state
of the particle is modified by a parametric drive. Both the parametric drive and the feedback are turned off and -- as in the first part -- the
motion of the particle towards an equilibrated state is observerd.}
        \label{fig:setup}
    \end{center}
\end{figure}
On the left hand side we have the friction coefficient $\Gamma_0$ and the angular frequency that describes the fluctuation along the chosen axis. On
the right hand side, there are two forces. The first one is $F_\text{fluct}$, which describes a stochastic force caused by interactions with the gas
in the vacuum chamber. This force is given by
\begin{equation}
    F_\text{fluct} = \sqrt{2m\Gamma_0k_BT_0} \ \xi\left(t\right)
\end{equation}
where $T_0$ is the temperature of the heat bath (i.e. the surrounding gas in the vacuum chamber), $k_B$ is the Boltzmann constant and $\xi(t)$ is
white noise, which obeys the equations $\left\langle\xi(t)\right\rangle= 0$ and $\left\langle\xi(t)\xi(t')\right\rangle= \delta(t-t')$, which means
that it is a random force. The term $\Gamma_0$ appears in the formulat due to the fluctuation-dissipation theorem, which links the damping rate to the
stochastic force.\\
The external force $F_\text{ext}$ is part of the experimental setup, where the frequency of the fluctuation along an axis, $\Omega_0$, is measured and
used to suppress the motion along said axis. This causes a decrease in the particle fluctuations and thus acts as a cooling mechanism for the particle
in the trap. This process creates a non-equilibrium steady state $\rho_{ss}(u,\alpha)$, which is not known analytically. This state is the starting
point of this thesis.\\

\newpage
\section{Simulation}
%simulation techniques and general concept of the simulation itself, used methods and so on 
The problem at hand can be studied on an atomic level with the use of computer simulation. There is a variety of methods for computer simulations
that are widely used, one of which being Molecular Dynamics (MD) simulations. The following section will give a brief overview of the concepts of this
method, which is followed by the application to the simulation of the experiment.

\subsection{Molecular Dynamics}
Molecular Dynamics\cite{Frenkel2001} simulations is a technique for simulating, as the name suggests, the dynamics of a classical many-body system. In this case,
classical means, that the trajectories of the individual particles are calculated using classical mechanics rather then quantum mechanics. For
relatively big atoms/molecules this is a very good approximation, whereas for systems consisting of hydrogen or helium the effects of quantum
mechanics cannot be neglected and other methods (such as ab-initio simulation) has to be used.\\
The dynamics of the system are obtained by solving Newton's equations of motion for every particle.

\subsection{The Glass Particle}

@@

Since $\varepsilon$ and $\sigma$ are crucial parameters for the simulation and do (usually) not change over time, it is practical to use them to

@@

Since the reduced units will be used throughout the rest of this thesis, i will drop the asterisk henceforth.\\

@@

With the choice of FCC as initial configuration, there are optimal numbers for the numbers of the particles in the system. Since once FCC cell (as
depicted) shares its atoms with its next neighbours, the number of atom per unit cell is 4, as depicted in fig. \ref{fig:wiki}. The whole system of
atoms is then created by repeating this cell structure. One convenient way is to arrange the unit cells in a cubic system, so if there are $M$ FCC
unit cells on one edge, the whole system consists of $M^3$ cells. Since there are 4 particles per cell, there are ideal or so called \textit{magic
numbers} for atoms for which this setup works perfectly: $N = 4M^3 = 4,32,108,256,500,864,\ldots$.\\

@@

        \begin{subfigure}[t]{0.3\textwidth}

@@

        \begin{subfigure}[t]{0.3\textwidth}
            \includegraphics[scale=0.15]{images/fcc_unit.png}
            \caption{Schematic of the number of particles contained in one FCC unit cell. The atoms on the edges are shared by 8 cells and the ones 
                on the faces by 2 cells\cite{unitcell}}
            \label{fig:fcccut}

@@

    p_2 &=& \{0.5,0.5,0\}\\
    p_3 &=& \{0.5,0,0.5\}\\
    p_4 &=& \{0,0.5,0.5\}

@@

From the particle number $N$ and the number of FCC unit cells per edge $M$ the lattice constant $a$ can be calculated

@@

When we look at the system from a thermodynamical standpoint, we see that it follows some kind of path in the phase space as time progresses. Every

@@

from one state to another. As mentioned above, this evolution (the dynamics of the system) is a crucial element to Molecular Dynamics. Since the

@@

expanding the coordinate mathbftor $\mathbf{r}_i$ for one particle after one time step $\Delta t$:

@@

Since the first derivative of the coordinate mathbftor is the velocity, $\dot{\mathbf{r}}_i(t) = \mathbf{v}_i(t)$, and the second derivative of the coordinate 
mathbftor is the acceleration, $\ddot{\mathbf{r}}_i(t) = \mathbf{a}_i(t)$, using Newton's law $\mathbf{F}_i(t) = m_i \mathbf{a}_i(t)$ equation \eqref{eq:verlet} 

@@

a combination of the Taylor expansions of the coordinate mathbftors at $t+\Delta t$ and $t-\Delta t$:

@@

@@

This algorithm is self starting, uses a small amount of memory, conserves the energy and gives a very good approximation to the originial trajectory

@@

    \item Calculate all the forces between the particles (for the first run only)

@@

The first point only has to be carried out for the first run, because the forces have not been calculated at this point. As the forces are calculated

@@

\subsection{The Laser Beam - Energy influx}
The glass sphere is trapped in the laser beam. While the motion of the center of mass is suppressed, which acts as a cooling mechanism, the

@@

To simulate this kind behaviour, thermostat algorithms \cite{Huenenberger2005} such as the Nos\'e-Hoover \cite{Nose1984,Hoover1985} or the Berendsen
\cite{Berendsen1984} algorithms.\\

@@

algorithm introduces an energy shift for longer simulation times. This led Wirnsberger et al \cite{Wirnsberger2015} to revisit the algorithm and 

@@

The simulation box, denoted by $\Omega$, and the regions $\Gamma_k$ are assumed to be moving with velocities $v_\Omega$ and $v_{\Gamma_k}$
respectively.\\

@@

This algorithm can be applied to the problem of the levitating nano sphere in the laser beam by adjusting the setup and some of the parameters.\\

@@

notation of fig. \ref{fig:ehex}, $\Omega = \Gamma_1$. Furthermore, neither the simulation box nor the heat source are moving, i.e. 

@@

force $\boldsymbol{\eta}$ and the correction term $\mathcal{E}\mathbf{r}$. Since there is only one region acting as a heat sink, the terms $\Delta Q$,

@@

This means that the laser is modeled to be pumping energy into the system, which increases the velocities of the velocities of the particles in the

@@

\subsection{The Laser Beam - Trapping}
As mentioned above, the glass particle is trapped in the laser beam and the motion of the particle is suppressed in every direction. This behaviour
has to be modeled as well.\\
The approximation of the original paper \cite{Gieseler2014}, where the movements in the three spatial directions are decoupled will be
used in the model as well. As to the model of the trapping force itself, the most straightforward approach is to use a harmonic oscillator potential.
The force and the corresponding potential can be written as
\begin{eqnarray}
    \mathbf{F} &=& -k x\\
    U &=& \frac12 k x^2
\end{eqnarray}
The trap is implemented by using the velocity Verlet algorithm with the above force to propagate the glass particle. Since the trap is acting on the
whole system, the force is acting on the center of mass position and velocity. To calculate the center of mass positions and velocities, the positions
and velocities of all particles have to be summed up and divided by the particle number:
\begin{eqnarray}
    \mathbf{r}_\text{COM} &=& \sum_{i=1}^N \frac{\mathbf{r}_i}{N}\\
    \mathbf{v}_\text{COM} &=& \sum_{i=1}^N \frac{\mathbf{v}_i}{N}
\end{eqnarray}
The position of the center of mass and the center of mass velocity are then propagated by using the same procedure as introduced in
section~\ref{section:verlet}.

\subsection{Surrounding Gas - Barostat}
Without any equilibrating mechanism the setup described by now would lead to the system melting, which is not desirable. In the next step we will 
introduce the surrounding gas of the gas chamber that will absorb some of the energy in the system, leading to the final state.\\
Generally, pressure is introduced to the system by surrounding the object of interest (in this case the glass nano sphere) with a pressure medium.

@@

pressure medium. The particles of this pressure medium flow into the simulation from an outside volume, whose geometry is based on the form of the
object of interest (this will be referred to as the minimal volume of cells). This barostat also acts as a thermostat and thus is suitable for the 
application in this thesis.\\

@@

gas particles reaching the inside of the nano crystal increases, which is not desirable. So $\sigma$ should be chosen small enough to keep the force
calculations at a minimum and large enough for the particle to stay on the outside of the crystal.\\ 

@@

This algorithm is formulated in a general fashion, so that a wide range of crystals and geometries can be used. The setup used in this thesis is a
FCC lattice, which has a nice symmetry that is not influenced by the surrounding gas. The algorithm will therefore not be carried out using 
cell-lists and a minimal volume of cells that is changing over the course of the simulation, but rather a fixed setup of the volume surrounding the 
nano crystal.\\ 
Since the crystal itself is a cube a straighforward approach is surrounding it by a bigger cube. The distance between the cube face and the nano
crystal is chosen to be the side length of the nano crystal, so the setup will properly scale for different numbers of particles. 
The schematic setup for this is depicted in fig.\ref{fig:gascube}.

@@

        \includegraphics[scale=0.5]{images/gas_volume_ver2.pdf}
        \caption{Schematic setup for the barostat. The outer volume (blue) is one side length $L$ away from the nano crystal (black). 
                    The gas particles are created along the blue lines of the outer volume. Since the geometry of the nano crystal does not change, 

@@

\newpage
\section{Results}
% results with different parameters, graphs, etc.
Since the simulation consists of various parts that have to work together, we first need to check if every part itself works.\\
\subsection{The Crystal}
The first element of the simulation is the setup of the FCC lattice on which the particles are placed on. To check if this is done correctly, we can
use a visualisation and rendering software such as VMD\cite{Humphrey1996} or OVITO\cite{Stukowski2010}. The rendered image of the nano crystal 
can be seen in fig.~\ref{fig:glass}. The chosen perspective shows the structure of the FCC lattice, where 3 layers are forming 
where the atoms take the position on the gap between the atoms from the layer below.\\
\begin{figure}[h]
    \begin{center}
        \includegraphics[scale=0.2]{images/fcc_rendering_layer_newcoloring.png}
        \caption{A rendered image of the FCC crystal. The chosen color-coding along the (1,1,1) Miller-index 
                highlights the layers for this perspective. The perspective is
                chosen to see the layering of the FCC crystal with its 3 layer structure. The black box acts as a reference point to outline the 
                geometry of the system.}
            %Rendering of the model of the nano crystal before starting the simulation. 
            %The chosen perspective shows the 3 layers of the FCC crystal with a
                    %These layers consist of atoms
                    %taking the position above the gap between two of the atoms from the layer below and match the positions of the atoms 2 layers
                %below. The black box is the outline of the particles' positions and shows the geometry of the system.}
        \label{fig:glass}
    \end{center}
\end{figure}

@@

The simulation is taking place in a NVE ensemble from a thermodynamic point of view. This means that the particle number $N$, the volume $V$ and the

@@

The next adjustment is a small one that makes a difference in the long run. When choosing random velocities, the system not only tends to have a
resulting velocity vector that causes the whole system to float away, but also a resulting angular momentum that causes the system to rotate.
The angular momentum can be defined via the moment of inertia $\mathbf{I}$
    \label{eq:angularmomentum}\mathbf{L} &=& \mathbf{I}\boldsymbol{\omega}\\
    \boldsymbol{\omega} &=& \frac{\mathbf{r} \times \mathbf{v}}{r^2}
where $\boldsymbol{\omega}$ is the angular velocity of the system. Another way to define the angular momentum is given by
    \label{eq:angularmomentum2}
    \mathbf{L} = \mathbf{r} \times \mathbf{p}
where $\mathbf{p}$ is the momentum of the particle, which in this case is equal to its velocity since the mass is set to 1.The moment of inertia 
$\mathbf{I}$ for a filled cube is given by 
    \mathbf{I} = 
    \begin{pmatrix}
        \frac{mL^2}{6} & 0 & 0\\
        0 & \frac{mL^2}{6} & 0\\
        0 & 0 & \frac{mL^2}{6} 
    \end{pmatrix}
where $m$ will be later set to 1 and $L$ is the side length of the cube. To use this equations to suppress the rotation of the cube, we need to use
equation~\eqref{eq:angularmomentum} and rewrite it as
    \label{eq:omega}
    \boldsymbol{\omega} = \mathbf{I}^{-1}\mathbf{L}
where $\mathbf{I}^{-1}$ is the inverse matrix of $\mathbf{I}$. The calculation of the inverse matrix is fairly straightforward, as it only contains
values in the diagonal, which means that the inverse matrix can be calculated by simply inverting these values. The angular momentum for every
particle $\mathbf{L}_i$ in the above equation be calculated using eq.~\eqref{eq:angularmomentum2}, where the position is taken to be the distance
between the particle and the center of mass
    \Delta\mathbf{r}_i = \mathbf{r}_\text{COM} - \mathbf{r}_i
The values of the angular momentum for each paricle are added up and multiplied by the inverse of the moment of inertia, as described in
eq.~\eqref{eq:omega}. The final step is updating the velocities of the particles:
    \mathbf{v}_i^\text{new} = \mathbf{v}_i - \left(\boldsymbol{\omega} \times \Delta\mathbf{r}_i\right)

@@

 \label{eq:rescale1}   E &=& \sum_{i=1}^N \mathbf{v}_i^2 \\
 \label{eq:rescale2}   \lambda &=& \sqrt{\frac{3(N-1)T}{E}} \\

@@

\subsection{Barostat}

@@

While the eHEX algorithm is pumping energy into the nano crystal while the surrounding gas particles absorb excessive energy. This leads to the
development of a new, higher steady temperature of the nano crystal, which can be seen in fig.~\ref{fig:baro1}. The first simulations were performed
with the eHEX and the barostat algorithm starting at the same point. This lead to the question: does the new steady temperature depend on the time it
takes the gas particles to reach the system? The answer is: no. After applying the barostat algorithm first to fill the box with gas atoms and then
applying the eHEX algorithm it can be seen that the resulting temperature is the same, which can be seen in fig.~\ref{fig:baro2}.

@@

        \includegraphics[scale=0.4]{images/baro_temp_compare_other.pdf}
        \caption{Two versions of the application of the eEHX and barostat algorithm. The shorter run (red, x-axis on the bottom) represents 
            the simultaneous start of the eHEX and the barostat algorithm. In the longer run (blue, x-axis on the top) there are 20000 steps, where
        the barostat algorithm is applied at the same time as the velocity Verlet to give the system time to equilibrate and only after the eHEX
    algorithm is taking the place of the velocity Verlet. The mean value of the resulting temperature is the same, indicating that the final
temperature is based on the values of the parameters, such as pressure and ambient temperature, rather than geometry.}
As we are interested in the temperatures of the particles before and after the interaction with the crystal, we need some way to differentiate those

@@

nano crystal. Since the gas particles do not interact with each other, the total energy of the gas particles does not change until they reach the
interaction range of the nano crystal. There they undergo a change of energy due to the interaction with the nano crystal and then make their way away

>>>>

\newpage
\section{Motivation}
%introduction of the experiment, the problem and the idea
\section{The experiment}
The starting point of this thesis is an experiment conducted by Gieseler et al \cite{Gieseler2014}. It is an optical tweezer experiment, where the
motion of a glass nanoparticle in a laser trap was used to investigate the fluctuation theorem\cite{Crooks1999}.
\subsection{Experimental setup}
In the experiment, a silica nano particle with a radius of about 75 nm and mass of about $3 \times 10^{-18}$kg is trapped in a laser beam within a
vacuum chamber. The trapping of the silica nano particle is achieved by a gradient force of the
laser beam acting on the particle. The experimental setup is depicted in fig.~\ref{fig:setup}.\\
The particle fluctuates within the trap in all three spatial directions. These fluctuations can be approximated
such that they are decoupled, which means that they can be described by a one-dimensional Langevin equation:
    \ddot{x} + \Gamma_0 \dot{x} + \Omega^2_0x = \frac 1 m \left(F_\text{fluct} + F_\text{ext}\right)
%The experiment is set up as follows: a glass sphere with radius $r \approx 75$ nm and mass $m \approx 3 \times 10^{-18}$ kg is trapped in a laser beam
%in a vacuum chamber. The initial state of the glass particle is prepared by modulation of the laser, such that the oscillation of the glass particle
%is suppressed in every direction. This is done by measuring the oscillation of the particle in every direction (assuming that the oscillation in every
%direction is decoupled) and using this information as feedback for the laser. This experimental setup can be seen in figure \ref{fig:setup}.
\begin{figure}[H]
        \includegraphics[scale=0.3]{images/experimental_setup.jpg}
        \caption{Experimental setup of the optical tweezer experiment. A silica nano particle is trapped in a laser beam via gradient force in a
        vacuum. The feedback is used to cool down the particle and create a non-equilibrium steady state. In the first part of the experiment, the
    feedback is turned off and the motion of the particle towards an equilibrated state is observed. In the second part of the experiment, the steady state
of the particle is modified by a parametric drive. Both the parametric drive and the feedback are turned off and -- as in the first part -- the
motion of the particle towards an equilibrated state is observerd. Figure taken from \cite{Gieseler2014}.}
        \label{fig:setup}
On the left hand side we have the position $x$ (and its derivatives $\dot{x}$, the velocity and $\ddot{x}$, the acceleration), the friction 
coefficient $\Gamma_0$ and the angular frequency $\Omega_0$ that describes the fluctuation along the chosen axis. On
the right hand side, there are two forces. The first one is $F_\text{fluct}$, which describes a stochastic force caused by interactions with the gas
in the vacuum chamber. This force is given by
    \label{eq:ffluct}
    F_\text{fluct} = \sqrt{2m\Gamma_0k_BT_0} \ \xi\left(t\right)
where $T_0$ is the temperature of the heat bath (i.e., the surrounding gas in the vacuum chamber), $k_B$ is the Boltzmann constant and $\xi(t)$ is
white noise, which obeys the equations $\left\langle\xi(t)\right\rangle= 0$ and $\left\langle\xi(t)\xi(t')\right\rangle= \delta(t-t')$, which means
that it is a random force. The term $\Gamma_0$ appears in the formula \eqref{eq:ffluct} 
due to the fluctuation-dissipation theorem, which links the damping rate to the stochastic force.\\
%The external force $F_\text{ext}$ is part of the experimental setup, where the frequency of the fluctuation along an axis, $\Omega_0$, is measured and
%used to suppress the motion along said axis. This causes a decrease in the particle fluctuations and thus acts as a cooling mechanism for the particle
%in the trap. This process creates a non-equilibrium steady state $\rho_{ss}(u,\alpha)$, which is not known analytically. This state is the starting
%point of this thesis.\\

\newpage
\section{Simulation}
%simulation techniques and general concept of the simulation itself, used methods and so on 
The problem at hand can be studied on an atomic level with the use of computer simulation. There is a variety of methods for computer simulations
that are widely used, one of which being molecular dynamics (MD) simulations.\\ 
The goal is to simulate the experiment depicted in Fig.~\ref{fig:setup} as accurately as possible. To achieve this, the setup has to be broken down
into individual pieced that can be modeled by using existing methods. The nano particle will be approximated by a cube of particles, sitting on an FCC
lattice that interact via a Lennard-Jones potential. The laser will be broken down into its two main purposes, as it acts as a trapping device as well
as a heat source for the nano particle. Surrounding the nano particle will be a gas chamber, that acts as a thermostat in the simulation, that will be
modeled as a box surrounding the nano particle. This box is filled with gas particles that interact with the nano particle via a soft-sphere potential
and do not interact with one another.\\
The following section will give a brief overview of the concepts used to simulate every part of the experiment.

\subsection{Molecular Dynamics}
Molecular dynamics~\cite{Frenkel2001} simulations is a technique for simulating, as the name suggests, the dynamics of a classical many-body system. In this case,
classical means, that the trajectories of the individual particles are calculated using classical mechanics rather then quantum mechanics. For
relatively big atoms/molecules this is a very good approximation, whereas for systems consisting of hydrogen or helium the effects of quantum
mechanics cannot be neglected and other methods have to be used.\\
The dynamics of the system are obtained by solving Newton's equations of motion for every particle.

\subsection{The Nano Particle}

@@

Since $\varepsilon$ and $\sigma$ are crucial parameters for the simulation and do not change over time, it is practical to use them to

@@

Since the reduced units will be used throughout the rest of this thesis, I will drop the asterisk henceforth.\\

@@

With the choice of FCC as initial configuration, there are optimal numbers for the numbers of the particles in the system. Since once FCC cell 
shares its atoms with its next neighbours, the number of atom per unit cell is 4 -- 1/8 of a particle on eight corners and 1/2 of a particle on six faces.
The whole system of atoms is then created by repeating this cell structure. One convenient way is to arrange the unit cells in a cubic system, 
so if there are $M$ FCC unit cells on one edge, the whole system consists of $M^3$ cells. Since there are 4 particles per cell, 
there are ideal or so called \textit{magic numbers} for atoms for which this setup works perfectly: $N = 4M^3 = 4,32,108,256,500,864,\ldots$.\\

@@

        \begin{subfigure}[t]{0.4\textwidth}

@@

@@

    p_2 &=& \{0.5*a,0.5*a,0\}\\
    p_3 &=& \{0.5*a,0,0.5*a\}\\
    p_4 &=& \{0,0.5*a,0.5*a\}

@@

From the particle number $N$ and the number of FCC unit cells per edge $M$ the \textit{lattice constant} $a$ can be calculated

@@

When we look at the system from a microscopic standpoint, we see that it follows some kind of path in the phase space as time progresses. Every

@@

from one state to another. As mentioned above, this evolution (the dynamics of the system) is a crucial element to molecular dynamics. Since the

@@

expanding the coordinate vector $\mathbf{r}_i$ for one particle after one time step $\Delta t$:

@@

Since the first derivative of the coordinate is the velocity, $\dot{\mathbf{r}}_i(t) = \mathbf{v}_i(t)$, and the second derivative of the coordinate 
is the acceleration, $\ddot{\mathbf{r}}_i(t) = \mathbf{a}_i(t)$, using Newton's law $\mathbf{F}_i(t) = m_i \mathbf{a}_i(t)$ equation \eqref{eq:verlet} 

@@

a combination of the Taylor expansions of the coordinate vectors at $t+\Delta t$ and $t-\Delta t$:

@@

        \mathbf{v}_i(t+\Delta t) = \mathbf{v}_i(t) + \frac1{2m} \Big[\mathbf{F}_i(t) + \mathbf{F}_i(t+\Delta t)\Big] \Delta t

@@

This algorithm is self starting, uses a small amount of memory, conserves the energy (not exactly) 
and gives a very good approximation to the originial trajectory

@@

    \item Calculate all the forces between the particles (for the first step only)

@@

The first point only has to be carried out for the first step, because the forces have not been calculated at this point. As the forces are calculated

@@

\subsection{The Laser Beam - Energy Influx}
The nano particle is trapped in the laser beam. While the motion of the center of mass is localized, the

@@

To simulate this kind of behaviour, thermostat algorithms \cite{Huenenberger2005} such as the Nos\'e-Hoover \cite{Nose1984,Hoover1985} or the Andersen
\cite{Andersen1980} algorithms are often used in simulation to change the temperature of the system in a controllable way.\\

@@

algorithm introduces an energy shift for longer simulation times. This led Wirnsberger et al. \cite{Wirnsberger2015} to revisit the algorithm and 

@@

The centers of mass of the particles in simulation box, denoted by $\Omega$, and the regions $\Gamma_k$ 
are assumed to be moving with velocities $v_\Omega$ and $v_{\Gamma_k}$ respectively.\\

@@

This algorithm can be adapted for the problem of the levitating nano sphere in the laser beam by adjusting the setup and some of the parameters.\\

@@

notation of fig. \ref{fig:ehex}, $\Omega = \Gamma_1$. Furthermore, neither center of mass of the particles the simulation box nor 
the heat source are moving, i.e. 

@@

force $\boldsymbol{\eta}$ and the correction term $\mathcal{E}\mathbf{r}$. Since there is only one region acting as a heat source, the terms $\Delta Q$,

@@

This means that the laser is modeled to be pumping energy into the system, which increases the velocities of the particles in the

@@

\subsection{The Laser Beam - Trapping}
As mentioned above, the glass particle is trapped in the laser beam and the position of the particle is localized. This behaviour
has to be modeled as well.\\
The approximation of the original paper \cite{Gieseler2014}, where the movements in the three spatial directions are decoupled will be
used in the model as well. As to the model of the trapping force itself, the most straightforward approach is to use a harmonic oscillator potential.
The force and the corresponding potential can be written as
\begin{eqnarray}
    \mathbf{F} &=& -k \Big[\mathbf{x}-\mathbf{x}_0\Big]\\
    U &=& \frac12 k \Big[\mathbf{x}-\mathbf{x}_0\Big]^2
\end{eqnarray}
where $\mathbf{x}_0$ is the position of minimal potential energy. 
Since the trap is acting on the whole system, the force is acting on the center of mass. 
To calculate the center of mass positions and velocities, the positions and velocities of all particles have to be summed up and 
divided by the particle number:
\begin{equation}
    \mathbf{r}_\text{COM} = \frac1N\sum_{i=1}^N \mathbf{r}_i
\end{equation}

\subsection{Surrounding Gas - Thermostat}
Without any equilibrating mechanism the setup described by now would lead to the system heating up indefinetly, which is not desirable. Furthermore,
the goal is to recreate the experiment as close as possible. Since the nano particle is trapped in a laser beam within a vacuum chamber, the
surrounding gas has to be modeled as well. In the next step we will introduce the surrounding gas of the gas chamber that will absorb some of the 
energy in the system, leading to the final state.\\
Generally, pressure is introduced to the system by surrounding the object of interest (in this case the glass nano sphere) with a thermostatting
pressure medium.

@@

thermostatting pressure medium. The particles of this pressure medium flow into the simulation from an outside volume, whose geometry is based on the form of the
object of interest (this will be referred to as the minimal volume of cells) and leave the simulation box if their position reaches the boundary of
this minimal volume of cells. This behaviour is very close to the real experiment which makes this thermostat an ideal candidate for the simulation of
the experiment. 
%This barostat also acts as a thermostat and thus is suitable for the 
%application in this thesis.\\

@@

gas particles reaching the inside of the nano particle increases, which is not desirable. So $\sigma$ should be chosen small enough to keep the force
calculations at a minimum and large enough for the particle to stay on the outside of the crystal. In the computer simulations performed for this
thesis, the interaction length is chosen to be $\sigma = 1$ with a cut-off radius of $r_c = 2.5$.\\ 

@@

This algorithm is formulated in a general fashion, so that a wide range of crystals and geometries can be used. 
%The setup used in this thesis is a FCC lattice, which has a nice symmetry that is not influenced by the surrounding gas. 
Although the shape of the nano particle changes over the course of the simulation, it is not necessary to change the geometry of the surrounding box.
The algorithm will therefore not be carried out using cell-lists and a minimal volume of cells that is changing over the course of the simulation, 
but rather a fixed setup of the volume surrounding the nano particle.\\ 
Since the particle itself is modeled as a cube, a straighforward approach is surrounding it by a bigger cube. The distance between the cube face and the nano
particle is chosen to be the side length of the nano particle, so the setup will properly scale for different numbers of particles. 
The schematic setup for this is depicted in Fig.\ref{fig:gascube}.

@@

        \includegraphics[scale=0.5]{images/gas_volume_ver3.pdf}
        \caption{Schematic setup for the barostat. The outer volume (blue) is one side length $s$ away from the nano particle (black) which means that
            the side length of the face of the surrounding box is $L=3s$. 
                    The gas particles are created along the blue lines of the outer volume. Since the overall geometry of the nano particle does 
                    not change, 

@@

\newpage
\section{Results}
% results with different parameters, graphs, etc.
Since the simulation consists of various parts that have to work together, we first need to check if every part itself works.\\
\subsection{The Crystal}
The first element of the simulation is the setup of the FCC lattice on which the particles are placed on. To check if this is done correctly, we can
use a visualisation and rendering software such as VMD\cite{Humphrey1996} or OVITO\cite{Stukowski2010}. The rendered image of the nano particle 
can be seen in fig.~\ref{fig:glass}. The chosen perspective shows the structure of the FCC lattice, where 3 layers are forming 
where the atoms take the position on the gap between the atoms from the layer below.\\
\begin{figure}[h]
    \begin{center}
        \includegraphics[scale=0.2]{images/fcc_rendering_layer_newcoloring.png}
        \caption{A rendered image of the FCC lattice. The chosen color-coding along the (1,1,1) Miller-index 
                highlights the layers for this perspective. The perspective is
                chosen to see the layering of the FCC lattice with its 3 layer structure. The black box acts as a reference point to outline the 
                geometry of the system.}
            %Rendering of the model of the nano crystal before starting the simulation. 
            %The chosen perspective shows the 3 layers of the FCC crystal with a
                    %These layers consist of atoms
                    %taking the position above the gap between two of the atoms from the layer below and match the positions of the atoms 2 layers
                %below. The black box is the outline of the particles' positions and shows the geometry of the system.}
        \label{fig:glass}
    \end{center}
\end{figure}

@@

The velocity Verlet part of the simulation is taking place in a NVE ensemble from a thermodynamic point of view. 
This means that the particle number $N$, the volume $V$ and the

@@

%The next adjustment is a small one that makes a difference in the long run. When choosing random velocities, the system not only tends to have a
%resulting velocity vector that causes the whole system to float away, but also a resulting angular momentum that causes the system to rotate.
%The angular momentum can be defined via the moment of inertia $\mathbf{I}$
%\begin{eqnarray}
    %\label{eq:angularmomentum}\mathbf{L} &=& \mathbf{I}\boldsymbol{\omega}\\
    %\boldsymbol{\omega} &=& \frac{\mathbf{r} \times \mathbf{v}}{r^2}
%\end{eqnarray}
%where $\boldsymbol{\omega}$ is the angular velocity of the system. Another way to define the angular momentum is given by
%\begin{equation}
    %\label{eq:angularmomentum2}
    %\mathbf{L} = \mathbf{r} \times \mathbf{p}
%\end{equation}
%where $\mathbf{p}$ is the momentum of the particle, which in this case is equal to its velocity since the mass is set to 1.The moment of inertia 
%$\mathbf{I}$ for a filled cube is given by 
%\begin{equation}
    %\mathbf{I} = 
    %\begin{pmatrix}
        %\frac{mL^2}{6} & 0 & 0\\
        %0 & \frac{mL^2}{6} & 0\\
        %0 & 0 & \frac{mL^2}{6} 
    %\end{pmatrix}
%\end{equation}
%where $m$ will be later set to 1 and $L$ is the side length of the cube. To use this equations to suppress the rotation of the cube, we need to use
%equation~\eqref{eq:angularmomentum} and rewrite it as
%\begin{equation}
    %\label{eq:omega}
    %\boldsymbol{\omega} = \mathbf{I}^{-1}\mathbf{L}
%\end{equation}
%where $\mathbf{I}^{-1}$ is the inverse matrix of $\mathbf{I}$. The calculation of the inverse matrix is fairly straightforward, as it only contains
%values in the diagonal, which means that the inverse matrix can be calculated by simply inverting these values. The angular momentum for every
%particle $\mathbf{L}_i$ in the above equation be calculated using eq.~\eqref{eq:angularmomentum2}, where the position is taken to be the distance
%between the particle and the center of mass
%\begin{equation}
    %\Delta\mathbf{r}_i = \mathbf{r}_\text{COM} - \mathbf{r}_i
%\end{equation}
%The values of the angular momentum for each paricle are added up and multiplied by the inverse of the moment of inertia, as described in
%eq.~\eqref{eq:omega}. The final step is updating the velocities of the particles:
%\begin{equation}
    %\mathbf{v}_i^\text{new} = \mathbf{v}_i - \left(\boldsymbol{\omega} \times \Delta\mathbf{r}_i\right)
%\end{equation}

@@

\label{eq:rescale1}   E &=& \sum_{i=1}^N \frac{\mathbf{v}_i^2}{2} \\
 \label{eq:rescale2}   \lambda &=& \sqrt{\frac{3(N-1)T}{2E}} \\

@@

\subsection{Thermostat}

@@

While the eHEX algorithm is pumping energy into the nano particle while the surrounding gas particles absorb excessive energy. This leads to the
development of a new, higher steady temperature of the nano particle, which can be seen in Fig.~\ref{fig:baro1}. 
%The first simulations were performed
%with the eHEX and the barostat algorithm starting at the same point. This lead to the question: does the new steady temperature depend on the time it
%takes the gas particles to reach the system? The answer is: no. After applying the barostat algorithm first to fill the box with gas atoms and then
%applying the eHEX algorithm it can be seen that the resulting temperature is the same, which can be seen in fig.~\ref{fig:baro2}.

@@

%\begin{figure}[h]
        %\includegraphics[scale=0.4]{images/baro_temp_compare_other.pdf}
        %\caption{Two versions of the application of the eEHX and barostat algorithm. The shorter run (red, x-axis on the bottom) represents 
            %the simultaneous start of the eHEX and the barostat algorithm. In the longer run (blue, x-axis on the top) there are 20000 steps, where
        %the barostat algorithm is applied at the same time as the velocity Verlet to give the system time to equilibrate and only after the eHEX
    %algorithm is taking the place of the velocity Verlet. The mean value of the resulting temperature is the same, indicating that the final
%temperature is based on the values of the parameters, such as pressure and ambient temperature, rather than geometry.}
        %%\caption{Temperature of the system during the application of the eHEX and the barostat algorithm. During the first 20000 steps the system is
        %%in equilibrium while the velocity Verlet algorithm is applied. After that, the eHEX and the barostat algorithms are turned on. This leads to
    %%an increase in temperature until the gas particles reach the nano crystal to slow down the heating of the system. After some time, a plateau is
%%reached and the temperature seems to stay at this value.}
        %\label{fig:baro1}
As we are interested in the temperatures of the particles before and after the interaction with the nano particle, we need some way to differentiate those

@@

nano particle. Since the gas particles do not interact with each other, the total energy of the gas particles does not change until they reach the
interaction range of the nano particle. There they undergo a change of energy due to the interaction with the nano particle and then make their way away
from the nano particle. After the interaction, when they reach a point outside of the interaction range, the enrgy stays constant again until the
